A Sample {\ttlit ACM} SIG Proceedings Paper in LaTeX Format\titlenote{(Does NOT produce the permission block, copyright information nor page numbering). For use with ACM\_PROC\_ARTICLE-SP.CLS. Supported by ACM.}

### 2014-03-07 paper/draft.tex
First Revision

The second class of issues concerns platform integrity and user data privacy. On one hand, from an user-centric point of view, the user would like to preserve privacy of its data when it stores data on the cloud platform. However, hypervisor is normally the most privileged task executed in the platform and therefore a curious or malicious administrator could easily retrieve those data \cite{cloudvisor:zhang}. Moreover, in multi-tenancy cloud platform, a user shares the same physical resources with other users. Several works \ref{sec:you,bot:att} showed the potential threat and risks of attacks issued by other users located on the same physical resource.

On the other hand, from a user-centric point of view, it is difficult to observe and inspect deeply the functioning of the whole platform, leveraging commodity general purpose hypervisors (GPH).General Purpose Hypervisors are executed at the most privileged level and so, it is complicated to detect and intervene on them in presence of malicious attack or misbehavior, without service interruption for all VMs hosted on the server. 
To sum up, both visions collide with the execution of the hypervisor at the highest level of privileged. This attitude prevents the platform administrator from a real monitoring of the whole platform, but also the possibility for users to concretely inspects actions that platform administrators do on user data. In the following paragraphs, we will make reference to this class of problem as ``Vertical'' requirements.

To resume, in order to present a complete view of the architecture, we identified five key design principles, that a cloud computing platform has to implement in order to solve the cited problems. In the following paragraphs, those principles will represent a grid of analysis of different architectures.
\begin{itemize}[]
\item \textbf{Inter-platform support} The cloud architecture has to be designed in order to support directly different platforms and service providers. The same user virtual instances have to be executed ( and potentially migrated) unmodified over different platforms.
\item \textbf{Fail-Safe modular architecture} In order to recover soft failures, the archtecure is supposed to be fail-safe, being able of automatically migrate instances or user-data in presence of a menace, but also reactivating the correct functioning in presence of a software failure. In order to satisfy this property, the amount of safety-critical components has to be as lowest as possible and the overall design has to be kept as modular as possible
\item \textbf{Small TCB} The amount of lines of code executed at the highest level of privilege is normally proportional to the number of potential failures and bugs in the platform. According to this, an architecture presenting a tiny TCB will improve its safety and integrity properties by design.
\item \textbf{User-Data privacy} A curious or malicious administrator ( or another platform user) would not have the right to inspect and analyze user data and instances, without the explicit consensus with the user above mutually trusted services. User-data privacy is a major challenge of modern cloud platforms. It is important to identify a trade-off that allows providers to identify and intervene on possible menaces, but also guarantee the impossibility for a curious or malicious operator to access to them without restrictions. In order to respect this principle when a platform administrator manages the platform it has not the highest level of privilege and each action on users object that he wants to deliver has to be validated by a mutually-trusted core, executed at the highest level of privilege. Moreover, this mutually-trust core has also to apply memory page cipher/uncipher in order to prevent memory page snooping when a page is swapped out.
\item \textbf{3rd part compatibility} Avoid disruption of existing 3rd part applications compatibility could represent a major principle to consider. Platform interfaces has to be kept as close as possible to the existing one, preventing application rewrite.
\end{itemize}
Moreover, beside this functional design principles specifying which features have to be achieved by the platform, another set of non-design principles could be identified. Those principles does not concern directly features or aspects of the architecture of the platform, but the effective feasibility of the architecture. 
\begin{itemize}[]
\item \textbf{Minimize code addition} a major issue could be represented by the amount of code lines that have to be added and maintained to an existing architecture to meet the functional requirement. Even it does not represent strictly a technical limitation to a certain proposal, it could become a ``show-stopper'' factor for its utilization.
\end{itemize}
In our analysis, it will be kept in consideration not only the overall suitability of the architecture, but also a quantification of the effort needed to develop and deploy it. In table \ref{int:des} it is possible to observe a resume of previous cited principles.

@@

\section{Related Work}
% Intro + performance issues
Cloud computing feature set could be enlarged exploiting two layers of virtualization instead of just one. The presence of two different hypervisors allows the possibility to specialize hypervisors functions. Instead of having a general purpose hypervisor we could split the hypervisor role in two different layers, interconnected with a well defined interface.
Each layer will have to care about a subset of features, simplifying overall development.
However, nested virtualization on x86 architecture faced impeding performance issues \cite{rec:virt}.

@@

Other approaches leveraging only one layer of virtualization does not achieve the whole design-principle satisfaction. Self-service cloud approach (SSC) \cite{ssc:art} design a complete security architecture to overcome major security and privacy issues of currently deployed cloud infrastructures, in order to alleviate the different needing user-centric and provider-centric vision of cloud computing presented in \ref{par:intro}. Even if this architecture reproduces an approach similar to the SuperCloud without leveraging an extra layer of virtualization (DP4), the approach changes the behavior where the user acts introducing new components that have to be directly managed by the user. Therefore, the main drawback of this solution is to change the interface with the user, preventing him to straightforwardly move to the new platform without having to change its applications (DP5).

@@

On the other hand, L1 is responsible for horizontal feature set. it has to provide the widest possible support to different platforms and the capacity to virtualize adopting different techniques ( para-virtualization, hardware-assisted, dynamic binary translation). In this sense, Williams et al. propose a patched version of Xen, Xen Blanket, to accomplish a L1 virtualization layer, designed to run over different platforms. This concretely allows to run over different physical platform the same VM, and live-migrate it at the occurrence.

@@

Secondly, due to the monolithic architecture, each feature introduced in GPH enlarges the TCB. Therefore, leveraging GPH, a huge amount of code is running in the most privileged mode and has to be trust. Several works discuss and analyze the overall TCB of modern GPH \cite{nova, xmhf}, considering GPH TCB always bigger than 100 KLoC. This approach directly collides with one design principles introduced in the introduction paragraph (DP3), that identified a small TCB as a major requirement for a solid architecture.
%A brief comparison of TCB of widespread GPH is depicted in figure \ref{fig:tcbchart}. 
%More importanly, this is particularly true looking at the monolithic approach of commercial hyper-visors, where any addition to the hypervisor body enlarges the TCB.

@@

Microkernels have the capacity of easily redesign the hierarchy of privileges. A user application has normally to trust the micro-kernel and the set of user-space software it uses. However, this set of trusted set of drivers and tools have not to be trusted by every application executed in the system. Each different user application could use its own set of tools, correctly registering and configuring them. Normally, this allows microkernels to implement different ``personalities''. User applications shares the system but could have to deal with completely different system services in a transparent way. However, ``personalities'' interest had been hugely reduced by the fact that system services, device drivers and even application had to be completely re-conceived to fit this new design pattern. Therefore, with the introduction of VMs, microkernels lost the battle in favor of hypervisors, in terms of popularity. However, the great advantage of tiny TCB, provided by microkernel, was not provided also by hypervisors. As we analyzed before, they present normally a monolithic architecture, like the great part of modern OS. This feature kept alive the interest in microkernels, in particular trying to couple microkernel benefits with virtualization.

Therefore, leveraging this concept of microkernels, a micro hypervisor (MH) is composed from a small kernel providing only essential functions, like scheduling, IPC, memory management and VMX extension. Several micro-hypervisor have been proposed in literature \cite{xmhf, nova}. They rely on the micro-kernel paradigm, extending it to virtualization. MH naturally satisfy vertical design principles (DP1,DP3, DP4). The tiny core could be considered the only system-critical componente of the architecture and if it possible to assure its correctness it could represent the last line of defense, where it could be possible to rebuild the architecture.

@@

Finally, for sake of keeping the TCB small, MH could generally lack of several virtualization features, commonly available on modern GPH hypervisor. For instance, XMHF does not support multiple guests. This kind of drawbacks could limits the potential interest in MH, even they are able to introduce several important features.

@@

The first analyzed architecture concerns only GPH. At first, it is important to recall the most interesting achievable by such architectures. In particular, several remarkable features are live migration over different platform, transparent migration of virtualized clusters between different data-centers and improved isolation between customer instances and administration, improved control of the nested virtualization layer L1 (horizontal requirements DP1).
Supposing the completeness and the correctness of nested virtualization hardware facilities in widespread hypervisor, GPH over GPH represents the best achievable architecture to respect horizontal design principles, but it lacks in vertical DPs as the size of TCB is the same of currently deployed architectures.

@@

To sum up,  the adoption of MH could  increase feature set introduced by nested virtualization adoption. The first impeding problem is the implementation of MH, able to propose to guest to virtualize with hardware assistance.

@@

\begin{itemize}[]
\item \textbf{Inter-platform features} Executed as L1, Xen would virtualize its nested guest initially using PV delivering the compatibility layer.
\item \textbf{Tiny TCB} The NOVA kernel will represent the only part of the platform executed at the highest level of privilege. Even the VMM, Vancouver, will be executed in the L0 user-space. The micro-kernel in L0 enforces a strictly modular approach to L0. Therefore, the VMM will present distinct processes instances for each L1 guest, enforcing a context division.
\item \textbf{3rd part compatibility} From a user point of view the architecture still exports the same interfaces, allowing the user to keep its applications unchanged.
\item \textbf{Inter-platform support}
\item \textbf{Fail-safe components architecture}  Moreover, the tiny core is the only safety-critical software component. The system could potentially be able to recover any crash in different user-space components.
\end{itemize}

>>>>

The second class of issues concerns platform integrity and user data privacy. On one hand, from an user-centric point of view, the user would like to preserve privacy of its data when it stores data on the cloud platform. However, hypervisor is normally the most privileged task executed in the platform and therefore a curious or malicious administrator could easily retrieve those data \cite{cloudvisor:zhang}. Moreover, in multi-tenancy cloud platform, a user shares the same physical resources with other users. Several works \cite{sec:you, bot:att} showed the potential threat and risks of attacks issued by other users located on the same physical resource.
On the other hand, from a user-centric point of view, it is difficult to observe and inspect deeply the functioning of the whole platform, leveraging commodity general purpose hypervisors (GPHs).GPHs are executed at the most privileged level and so, it is complicated to detect and intervene on them in presence of malicious attack or misbehavior, without service interruption for all VMs hosted on the server. 
To sum up, both visions collide with the execution of the hypervisor at the highest level of privileged. This attitude prevents the platform administrator from a real monitoring of the whole platform, but also the possibility for users to concretely inspects actions that platform administrators do on user data. A major challenge of new generation cloud platform is to alleviate this different tension, finding an acceptable trade-off. In the following paragraphs, we will make reference to this class of problem as ``Vertical'' requirements.

To resume, in order to present a complete view of the architecture, we identified five key design principles, that a cloud computing platform has to implement in order to solve the cited problems. In the following paragraphs, those principles will represent a grid of analysis of different architectures.
\begin{itemize}[]
\item \textbf{DP1 User-Data privacy} A curious or malicious administrator ( or another platform user) would not have the right to inspect and analyze user data and instances, without the explicit consensus with the user above mutually trusted services. User-data privacy is a major challenge of modern cloud platforms. It is important to identify a trade-off that allows providers to identify and intervene on possible menaces, but also guarantee the impossibility for a curious or malicious operator to access to them without restrictions. In order to respect this principle when a platform administrator manages the platform it has not the highest level of privilege and each action on users object that he wants to deliver has to be validated by a mutually-trusted core, executed at the highest level of privilege. Moreover, this mutually-trust core has also to apply memory page cipher/uncipher in order to prevent memory page snooping when a page is swapped out.
\item \textbf{DP2 Fail-Safe modular architecture} In order to recover soft failures, the archtecure is supposed to be fail-safe, being able of automatically migrate instances or user-data in presence of a menace, but also reactivating the correct functioning in presence of a software failure. In order to satisfy this property, the amount of safety-critical components has to be as lowest as possible and the overall design has to be kept as modular as possible
\item \textbf{DP3 Small TCB} The amount of lines of code executed at the highest level of privilege is normally proportional to the number of potential failures and bugs in the platform. According to this, an architecture presenting a tiny TCB will improve its safety and integrity properties by design.
\item \textbf{DP4 Inter-platform support} The cloud architecture has to be designed in order to support directly different platforms and service providers. The same user virtual instances have to be executed ( and potentially migrated) unmodified over different platforms.
\item \textbf{DP5 3rd part compatibility} Avoid disruption of existing 3rd part applications compatibility could represent a major principle to consider. Platform interfaces has to be kept as close as possible to the existing one, preventing application rewrite.
\end{itemize}
Moreover, beside this functional design principles specifying which features have to be achieved by the platform, another set of non-design principles could be identified. Those principles does not concern directly features or aspects of the architecture of the platform, but the effective feasibility of the architecture. 
\begin{itemize}[]
\item \textbf{NDP1 Minimize code addition} a major issue could be represented by the amount of code lines that have to be added and maintained to an existing architecture to meet the functional requirement. Even it does not represent strictly a technical limitation to a certain proposal, it could become a ``show-stopper'' factor for its utilization.
\end{itemize}
In our analysis, it will be kept in consideration not only the overall suitability of the architecture, but also a quantification of the effort needed to develop and deploy it. In table \ref{int:des} it is possible to observe a resume of previous cited principles.

@@

\section{Related Work}
% Intro + performance issues
Several works presented in literature integrates these a subset of this design principles in their architecture, leveraging nested virtualization \cite{turtle:ibm,art:blan, cloudvisor:zhang}
Cloud computing feature set could be enlarged exploiting two layers of virtualization instead of just one. The presence of two different hypervisors allows the possibility to specialize hypervisors functions. Instead of having a general purpose hypervisor we could split the hypervisor role in two different layers, interconnected with a well defined interface.
Each layer will have to care about a subset of features, simplifying overall development.
However, nested virtualization on x86 architecture faced impeding performance issues \cite{rec:virt}.

@@

Other approaches leveraging only one layer of virtualization does not achieve the whole design-principle satisfaction. Self-service cloud approach (SSC) \cite{ssc:art} design a complete security architecture to overcome major security and privacy issues of currently deployed cloud infrastructures, in order to alleviate the different needing user-centric and provider-centric vision of cloud computing presented in \ref{par:intro}. Even if this architecture reproduces an approach similar to the SuperCloud without leveraging an extra layer of virtualization (DP4), the approach changes the behavior where the user acts introducing new components that have to be directly managed by the user. Therefore, the main drawback of this solution is to change the interface with the user, preventing the possibility to straightforwardly move to the new platform without having to change its applications (DP5).

@@

On the other hand, L1 is responsible for horizontal feature set. it has to provide the widest possible support to different platforms and the capacity to virtualize adopting different techniques ( para-virtualization, hardware-assisted, dynamic binary translation). According to these requirements, Williams et al. \cite{art:blan} propose a patched version of Xen, Xen Blanket, to accomplish a L1 virtualization layer, designed to run over different platforms. This concretely allows to run over different physical platform the same VM, and live-migrate it at the occurrence.

@@

Secondly, due to the monolithic architecture, each feature introduced in GPH enlarges the TCB. Therefore, leveraging GPH, a huge amount of code is running in the most privileged mode and has to be trust. Several works discuss and analyze the overall TCB of modern GPH \cite{nova, xmhf}, considering GPH TCB always bigger than 100 KLoC. This approach directly collides with one design principles introduced in the introduction paragraph (DP3), 
%that identified a small TCB as a major requirement for a solid architecture.
%A brief comparison of TCB of widespread GPH is depicted in figure \ref{fig:tcbchart}. 
%More importanly, this is particularly true looking at the monolithic approach of commercial hyper-visors, where any addition to the hypervisor body enlarges the TCB.

@@

Microkernels have the capacity of easily redesign the hierarchy of privileges. A user application has normally to trust the tiny micro-kernel and only the set of user-space software it uses. However, this set of trusted set of drivers and tools have not to be trusted by every application executed in the system. Normally, this allows microkernels to implement different ``personalities''. User applications shares the system but could have to deal with completely different system services in a transparent way. However, ``personalities'' interest had been hugely reduced by the fact that system services, device drivers and even application had to be completely re-conceived to fit this new design pattern. Therefore, with the introduction of VMs, microkernels lost the battle in favor of hypervisors, in terms of popularity. However, the great advantage of tiny TCB, provided by microkernel, was not provided also by hypervisors. As we analyzed before, they present normally a monolithic architecture, like the great part of modern OS. This feature kept alive the interest in microkernels, in particular trying to couple microkernel benefits with virtualization.

Therefore, leveraging this concept of microkernels, a micro hypervisor (MH) is composed from a small kernel providing only essential functions, like scheduling, IPC, memory management and VMX extension. Several micro-hypervisor have been proposed in literature \cite{xmhf, nova}. They rely on the micro-kernel paradigm, extending it to virtualization. MH naturally satisfy vertical design principles (DP1,DP3, DP4). The tiny core could be considered the only system-critical componente of the architecture and, due to its relative small dimension and the consequent minor attitude to vulnerabilities,it could represent the ``last line of defense''.

@@

Finally, microkernel modularization introduce a major challenge in architecture design. Due the increased complexity, for sake of kernel size, MH could generally lack of several virtualization features, commonly available on modern GPH hypervisor. For instance, XMHF does not support multiple guests. This kind of drawbacks could limits the potential interest in MH, even they are able to introduce several important features.

@@

The first analyzed architecture concerns only GPH. At first, it is important to recall the most interesting achievable by such architectures. In particular, several remarkable features are live migration over different platform, transparent migration of virtualized clusters between different data-centers (horizontal requirements DP1).
Supposing the completeness and the correctness of nested virtualization hardware facilities in widespread hypervisor, GPH over GPH satisfies all horizontal design principles, but it lacks in vertical DPs as the size of TCB is the same of currently deployed architectures.

@@

To sum up,  the adoption of MH could  increase feature set introduced by nested virtualization adoption. The first impeding problem is the implementation of MH, able to propose to guest to virtualize with hardware assistance. In next section, we present an implementation of GPH over MH, that try to circumvent this limitation leveraging paravirtualization.

@@

\begin{itemize}[]
\item \textbf{Inter-platform support} Executed as L1, Xen would virtualize its nested guest initially using PV delivering the compatibility layer.
\item \textbf{Tiny TCB} The NOVA kernel will represent the only part of the platform executed at the highest level of privilege. Even the VMM, Vancouver, will be executed in the L0 user-space. The micro-kernel in L0 enforces a strictly modular approach to L0. Therefore, the VMM will present distinct processes instances for each L1 guest, enforcing a context division.
\item \textbf{3rd part compatibility} From a user point of view the architecture still exports the same interfaces provided by the Xen hypervisor, allowing the user to keep its applications unchanged.
\item \textbf{Fail-safe components architecture}  Moreover, the tiny core is the only safety-critical software component. The system could potentially be able to recover any crash in different user-space components.
\end{itemize}
